---
title: "English Channel Geostrophic Wind Metrics"
---

The following plots show the sea-level pressure recorded in London,
Paris and De Bilt over the last year and the geostrophic wind that has
been derived from these three sites.

The plots are initially zoomed into the last 31 days but using the
slider below the plots the timespan can be expanded up to the a year.

These values are calculated from the SLP readings taken at 1200 UTC.

These values are updated on a daily basis using values distributed via
the Global Telecommunications System (GTS).

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE)
```


```{r, echo=FALSE, fig.show="hold", out.width="100%",warning=FALSE,message=FALSE}
library(dygraphs)
library(xts)
library(data.table)
library(lubridate)
library(knitr)
library(dplyr)
library(tidyr)

geowind_clim <- readRDS("../../../data/geowind_climatology.RDS") %>%
    rename(clim=val) %>%
    spread(quant,clim) %>%
    setnames(c("month","day",paste0("Q",c(0.05,0.25,0.5,0.75,0.95))))

geowind <- readRDS("../../../data/geowind_current.RDS") %>%
    filter(hour(time)==12,time>=now("UTC")-days(365)) %>%
    mutate(year=year(time),month=month(time),day=day(time))

geowind <- left_join(geowind,geowind_clim)

slp.xts <- xts(x=geowind[,c("debilt","london","paris")], order.by=geowind$time,tzone="UTC")
wind.xts <- xts(x=geowind[,c("geowind","Q0.5","Q0.05","Q0.95")], order.by=geowind$time,tzone="UTC")


date_range <- c(now("UTC")-months(3),now("UTC"))
dygraph(slp.xts, main = "MSLP",xlab="Date",ylab="MSLP (hPa)",group="geowind") %>%
dyRangeSelector(date_range)

dygraph(wind.xts, main = "Geostrophic Wind",xlab="Date",ylab="M/s") %>%
    dySeries(c("Q0.05","Q0.5","Q0.95"),"Median") %>%
    dyRangeSelector(date_range)
```
