---
authors: ["richard-cornes"]
categories: ["blog", "storms"]
date: '2023-11-03'
description: "The lowest barometric pressure in London since 1989"
tags: ["storms"]
title: "Storm Ciarán"
weight: 30
---

At 0800UTC on the 2nd November 2023, as storm Ciarán passed over
southern England, the barometric pressure recorded at Heathrow airport
in London sank to 955.7hPa.

Such low baraometric pressures are extremely rare and given the length
of the London pressure series, we are able to put that value in the context
of the last 300+ years.

The table below shows the dates of all sea-level pressures in the
series below 960hPa. These values represent the lowest values on a
given day, in order to give an indication of the low pressure values
associated with disrete storm events. This is not entirely reliable,
however, as can be seen in the double entry for the 1989 storm.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE)
```


```{r, echo=FALSE, fig.show="hold", out.width="100%",warning=FALSE,message=FALSE}
library(dplyr)
library(dataresqc)
library(lubridate)

## Read the data
data_dir <- "~/projects/geowindWEB/data/SEF"

ifiles <- list.files(data_dir,"London*",full=TRUE)

london <- lapply(ifiles,read_sef) %>%
    bind_rows()

london %>%
    filter(Value<960) %>%
    mutate(date=make_datetime(Year,Month,Day,Hour,Minute),
           date2=make_date(Year,Month,Day)) %>%
    group_by(date2) %>%
    arrange(Value) %>%
    slice(1) %>%
    ungroup() %>%
    select(date,Value) %>%
    arrange(Value) %>%
    as.data.frame() %>%
    knitr::kable(col.names=c("Time","Sea-level Pressure (hPa)"),align="rr") %>%
    kableExtra::kable_styling(full_width = FALSE)
```

As can see from this table the pressure on 2nd November 2023 was the
fourth lowest in the series, behind the 1989 and 1724 readings.
However, given the uncertainty with the 1724 measurements it likely to
be at least comparable to the pressure recorded at that time.

In Cornes (2014) I discussed the most significant storms to have
occurred across the British Isles since 1700. The low-pressure events
in 1724 were noted at the time both by observers in London and by the
French astromomer Giovanni Domenico Maraldi, who was recording
meteorological observations at the Paris Observatory. An observer
writing for the *British Journal* noted:

> *On Tuesday morning last [19 December 1724 NS] the quicksilver was
lower in all the barometers here than had ever been observed by our
most ancient Vertuosi[sic], even almost five tenths lower than in the
great November storm in 1703; and yet during the whole time of its
sinking, which was from Monday afternoon, between three or four of the
clock, to six the next morning, (in which space it fell an inch and
five tenths) we had no wind so considerable, as to do any remarkable
damage. Its greatest violence was from six to seven of the clock on
Monday night, when there fell a great quantity of rain, and the wind
(as far as I could perceive) was south east.*

As with Storm Ciarán, the strongest winds were not felt in southern
England but were experienced farther south. However, little is known
about that event and further research is needed.

Stephen Burt (University of Reading) has provided further details
about the low pressures recorded across Southern England on 2nd
November in this document:
https://www.reading.ac.uk/news/2023/Expert-Comment/Stephen-Burt-on-Storm-Ciaran

### References
Cornes, R.C. (2014), Historic storms of the northeast Atlantic since circa 1700: a brief review of recent research. Weather, 69: 121-125. https://doi.org/10.1002/wea.2289