---
authors: ["richard-cornes"]
categories: ["blog", "storms"]
date: '2023-11-01'
description: "The lowest barometric pressure in London since 1989"
tags: ["storms"]
title: "Storm Ciarán"
weight: 30
---


At 0800UTC on the 2nd November 2023, as storm Ciarán passed over southern England
the barometric pressure recorded at Heathrow airport in London sank to
955hPa.

Such events are extremely rare and given the length of the London
pressure, we are able to put that value in the context of the last
250+ years.

The table below shows the dates of all sea-level pressures in the
series below 960hPa. These values represent the lowest values on a
given day, as given the hourly sampling frequency for the 1989 event
there pressure was below 960hPa for most of the day.

As can see from this table the pressure on 2nd November 2023 was the
second lowest pressure in the series.

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse=TRUE)
```


```{r, echo=FALSE, fig.show="hold", out.width="100%",warning=FALSE,message=FALSE}
library(dplyr)
library(dataresqc)
library(lubridate)

## Read the data
data_dir <- "~/projects/geowindWEB/data/SEF"

ifiles <- list.files(data_dir,"London*",full=TRUE)

london <- lapply(ifiles,read_sef) %>%
    bind_rows()

london %>%
    filter(Value<960) %>%
    mutate(date=make_datetime(Year,Month,Day,Hour,Minute),
           date2=make_date(Year,Month,Day)) %>%
    group_by(date2) %>%
    arrange(Value) %>%
    slice(1) %>%
    ungroup() %>%
    select(date,Value) %>%
    arrange(Value) %>%
    as.data.frame() %>%
    knitr::kable(col.names=c("Time","MSLP"),align="rl")
```
In Cornes (2014) I discussed the most significant storms to have occurred across the British Isles since 1700. The 1724 event was noted at the time both by observers in London and by the French astromomer Maraldi, who was recording meteorological measurements at the Paris Observatory.

On the afternoon of 4th November the pressure is again forecast to drop below 960hPa in London, in certain weather forecast models/ensemble members. This table will therefore likely need updating again.



### References
Cornes, R.C. (2014), Historic storms of the northeast Atlantic since circa 1700: a brief review of recent research. Weather, 69: 121-125. https://doi.org/10.1002/wea.2289